---
title: "CAHOOTS Project - Analysis"
author: "Alex Gollier"
output: html_document
date: "2026-06-08"
---

```{r}
library(tidyverse)
library(readxl)
library(lubridate)

# Load cleaned data from cleaning.Rmd
eugene <- readRDS("eugene_cleaned.rds")
spd <- readRDS("spd_cleaned.rds")
mcslc <- readRDS("mcslc_cleaned.rds")
monthly_combined <- readRDS("monthly_combined.rds")
```


## Step 11: Time-Series 

```{r}
# Set CAHOOTS service stop date as reference line
cahoots_end <- as.Date("2025-04-01")

# Plot monthly call proportions by agency 
proportion_ts_plot <- monthly_combined %>%
  filter(city %in% c("Eugene", "Springfield")) %>%
  ggplot(aes(x = year_month, y = proportion, color = agency)) +
  geom_line() +
  facet_wrap(~ city, ncol = 1) +
  geom_vline(xintercept = cahoots_end, linetype = "dashed") +
  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = c(
    "EPD" = "#032C3C",
    "SPD" = "#2E86C1",
    "CAHOOTS" = "#BA351A",
    "MCSLC" = "#808080"
  )) +
  labs(
    title = "Monthly Call Proportions by Agency",
    x = "Month",
    y = "Proportion of Total Calls",
    color = "Agency"
  ) +
  theme_minimal()

# Display results
proportion_ts_plot

# Save as an image
ggsave("proportion_ts_plot.jpeg", plot = proportion_ts_plot, width = 10, height = 8, dpi = 300)
```

## Step 12: DiD Analysis

### 12.1: Overall Window Call Proportion DiD Analysis

```{r}
# Build regression dataset
did_reg_data <- monthly_combined %>%
  filter((city == "Eugene" & agency == "EPD") | (city == "Springfield" & agency == "SPD")) %>%
  mutate(treat = if_else(city == "Eugene", 1, 0), post = if_else(period == "after", 1, 0),
    treat_post = treat * post)

# Run DiD regression
did_model <- lm(proportion ~ treat + post + treat_post, data = did_reg_data)
summary(did_model)
```

### 12.2: Matched Calendar Window Call Proportion DiD Analysis

```{r}
# Filter to match the same calendar window 
did_reg_data_matched <- did_reg_data %>%
  filter((period == "after") |
    (period == "before" & year_month >= as.Date("2024-04-01") & year_month <= as.Date("2024-12-01"))
  )

# Rerun DiD with matched window
did_model_matched <- lm(proportion ~ treat + post + treat_post, data = did_reg_data_matched)

# Show Summary
summary(did_model_matched)
```


### 12.3: Overall Window Call Proportion DiD Analysis Plot

```{r}
# Build DiD summary data for plotting
did_data <- did_reg_data %>%
  group_by(city, period) %>%
  summarise(avg_proportion = mean(proportion), .groups = "drop")

# Build and save the DiD plot
did_plot <- did_data %>%
  mutate(period = factor(period,
                          levels = c("before", "after"),
                          labels = c("Before April 2025", "After April 2025"))) %>%
  ggplot(aes(x = period, y = avg_proportion, color = city, group = city)) +
  geom_line(linewidth = 2) +
  geom_point(size = 5) +
  geom_text(aes(label = scales::percent(avg_proportion, accuracy = 0.1),
                vjust = ifelse(city == "Springfield" & period == "After April 2025", 2.2, -1.5),
                hjust = ifelse(city == "Eugene" & period == "Before April 2025", 1.3, 0.5)),
            size = 4,
            show.legend = FALSE) +
  scale_y_continuous(limits = c(0.80, 0.97),
                     labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = c("Eugene" = "#BA351A", "Springfield" = "#032C3C")) +
  labs(
    title = "DiD Analysis: EPD vs SPD Call Share",
    x = NULL,
    y = "Proportion of Total Calls",
    color = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.text = element_text(size = 14),
    aspect.ratio = 1
  )

# Save as JPEG
ggsave("did_plot.jpeg", plot = did_plot, width = 6, height = 6, dpi = 300)

# Display
did_plot
```
### 12.4: Matched Window Call Proportion DiD Analysis Plot

```{r}
# Build matched-window summary data
did_data_matched <- did_reg_data_matched %>%
  group_by(city, period) %>%
  summarise(avg_proportion = mean(proportion), .groups = "drop")

# Build and save the matched-window DiD plot
did_plot_matched <- did_data_matched %>%
  mutate(period = factor(period,
                          levels = c("before", "after"),
                          labels = c("Before April 2025", "After April 2025"))) %>%
  ggplot(aes(x = period, y = avg_proportion, color = city, group = city)) +
  geom_line(linewidth = 2) +
  geom_point(size = 5) +
  geom_text(aes(label = scales::percent(avg_proportion, accuracy = 0.1),
                vjust = ifelse(city == "Springfield" & period == "After April 2025", 2.2, -1.5),
                hjust = ifelse(city == "Eugene" & period == "Before April 2025", 1.3, 0.5)),
            size = 4,
            show.legend = FALSE) +
  scale_y_continuous(limits = c(0.80, 0.97),
                     labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = c("Eugene" = "#BA351A", "Springfield" = "#032C3C")) +
  labs(
    title = "DiD Analysis: EPD vs SPD Call Share (Matched Window)",
    x = NULL,
    y = "Proportion of Total Calls",
    color = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.text = element_text(size = 14),
    aspect.ratio = 1
  )

# Save plot
ggsave("did_plot_matched.jpeg", plot = did_plot_matched, width = 6, height = 6, dpi = 300)

# Display
did_plot_matched
```
## Step 13: Parallel Trends Test

### 13.1: Overall Window Parallel Trends Test

```{r}
# Build pre-treatment EPD/SPD data 
pre_trend_data <- monthly_combined %>%
  filter((city == "Eugene" & agency == "EPD") |
         (city == "Springfield" & agency == "SPD")) %>%
  filter(period == "before") %>%
  mutate(
    treat = if_else(city == "Eugene", 1, 0),
    time_index = as.numeric(year_month)
  )

# Test for differential pre-trends
pre_trend_test <- lm(proportion ~ treat + time_index + treat:time_index, data = pre_trend_data)

# Show results
summary(pre_trend_test)
```

### 13.2: Overall Window Parallel Trends Chart

```{r}
parallel_trends_plot <- monthly_combined %>%
  filter((city == "Eugene" & agency == "EPD") |
           (city == "Springfield" & agency == "SPD")) %>%
  filter(period == "before") %>%
  ggplot(aes(x = year_month, y = proportion, color = city)) +
  geom_line() +
  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = c("Eugene" = "#BA351A", "Springfield" = "#032C3C")) +
  labs(
    title = "Parallel Trends Check: Pre-Treatment Period Only",
    x = NULL,
    y = "Proportion of Total Calls",
    color = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.text = element_text(size = 14),
    aspect.ratio = 0.5
  )

ggsave("parallel_trends_plot.jpeg", plot = parallel_trends_plot, width = 10, height = 5, dpi = 300)
parallel_trends_plot
```

### 13.3: Matched Window Parallel Trends Test

```{r}
# Parallel trends test for matched window 
pre_trend_data_matched <- monthly_combined %>%
  filter((city == "Eugene" & agency == "EPD") |
         (city == "Springfield" & agency == "SPD")) %>%
  filter(period == "before") %>%
  filter(year_month >= as.Date("2024-04-01") & year_month <= as.Date("2024-12-01")) %>%
  mutate(
    treat = if_else(city == "Eugene", 1, 0),
    time_index = as.numeric(year_month)
  )
pre_trend_test_matched <- lm(proportion ~ treat + time_index + treat:time_index,
                              data = pre_trend_data_matched)

# Show Results
summary(pre_trend_test_matched)
```


## 14. Volume-Based Analysis

### 14.1: Volume-Based Bar Chart

```{r}
# Build EPD volume across three time windows
epd_stability_matched <- eugene %>%
  filter(agency == "EPD") %>%
  mutate(
    period_label = case_when(
      calltime >= as.POSIXct("2025-04-08") ~ "Post Apr 2025",
      calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31") ~ "1-Year Pre",
      calltime >= as.POSIXct("2015-01-01") & calltime < as.POSIXct("2025-04-08") ~ "Pre Apr 2025",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(period_label)) %>%
  filter(as.character(year_month) != "2025-04-01") %>%
  filter(as.character(year_month) != "2023-01-01") %>%
  group_by(period_label, year_month) %>%
  summarise(monthly_count = n(), .groups = "drop") %>%
  group_by(period_label) %>%
  summarise(avg_monthly_calls = mean(monthly_count), .groups = "drop") %>%
  mutate(period_label = factor(period_label, levels = c(
    "Pre Apr 2025",
    "1-Year Pre",
    "Post Apr 2025"
  )))

# Plot
volume_plot <- epd_stability_matched %>%
  ggplot(aes(x = period_label, y = avg_monthly_calls, fill = period_label)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = scales::comma(round(avg_monthly_calls, 0))),
            vjust = -0.5, size = 4) +
  scale_y_continuous(limits = c(0, 8000),
                     labels = scales::comma) +
  scale_fill_manual(values = c(
    "Pre Apr 2025" = "#032C3C",
    "1-Year Pre" = "#032C3C",
    "Post Apr 2025" = "#BA351A"
  )) +
  labs(
    title = "EPD Average Monthly Call Volume Across Time Periods",
    x = NULL,
    y = "Average Monthly Calls",
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    legend.position = "none",
    axis.text.x = element_text(size = 14),
    aspect.ratio = 1
  )

# Save as JPEG
ggsave("volume_plot.jpeg", plot = volume_plot, width = 6, height = 6, dpi = 300)

# Display
volume_plot
```

### 14.2: Volume-Based Linear Regression

```{r}
# EPD volume change
epd_volume_data <- monthly_combined %>%
  filter(city == "Eugene" & agency == "EPD") %>%
  filter(
    (period == "after") |
    (period == "before" & year_month >= as.Date("2024-04-01") & year_month <= as.Date("2024-12-01"))
  ) %>%
  mutate(post = if_else(period == "after", 1, 0))

epd_volume_model <- lm(call_count ~ post, data = epd_volume_data)
summary(epd_volume_model)
```
## 15: Call-Type Analysis

### 15.1: CAHOOTS Call Type Frequency

```{r}
# Find the frequency of CAHOOTS call types in Eugene
eugene %>%
  filter(agency == "CAHOOTS", period == "before") %>%
  count(nature, sort = TRUE) %>%
  mutate(proportion = n / sum(n) * 100) %>%
  slice_head(n = 5) %>%
  print()
```

### 15.2: EPD Absorption Analysis

```{r}
# EPD before/after for top 4 CAHOOTS call types
top_cahoots_types <- c("PUBLIC ASSIST", "CHECK WELFARE", "TRANSPORT", "SUICIDAL SUBJECT")

# CAHOOTS average monthly calls per type — matched window
cahoots_monthly <- eugene %>%
  filter(agency == "CAHOOTS", period == "before") %>%
  filter(nature %in% top_cahoots_types) %>%
  filter(calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31")) %>%
  group_by(nature, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(nature) %>%
  summarise(CAHOOTS_Avg_Monthly = round(mean(n), 0), .groups = "drop")

# Create the table
epd_absorption_table <- eugene %>%
  filter(agency == "EPD") %>%
  filter(nature %in% top_cahoots_types) %>%
  filter(
    (period == "after") |
    (period == "before" & calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(nature, period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(nature, period) %>%
  summarise(avg_monthly = round(mean(n), 0), .groups = "drop") %>%
  pivot_wider(names_from = period, values_from = avg_monthly) %>%
  rename(EPD_Before = before, EPD_After = after) %>%
  mutate(
    Change = EPD_After - EPD_Before,
    Pct_Change = round((EPD_After - EPD_Before) / EPD_Before * 100, 1)
  ) %>%
  left_join(cahoots_monthly, by = "nature") %>%
  select(nature, CAHOOTS_Avg_Monthly, EPD_Before, EPD_After, Change, Pct_Change) %>%
  arrange(desc(CAHOOTS_Avg_Monthly))

# Show
print(epd_absorption_table)
```

### 15.3: EPD Absorption Plot

```{r}
# Build absorption data 
absorption_data <- eugene %>%
  filter(agency == "EPD") %>%
  filter(nature %in% c("PUBLIC ASSIST", "CHECK WELFARE", "SUICIDAL SUBJECT")) %>%
  filter(
    (period == "after") |
    (period == "before" & calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(nature, period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(nature, period) %>%
  summarise(avg_monthly = mean(n), .groups = "drop") %>%
  mutate(period = factor(period, levels = c("before", "after"),
                          labels = c("Pre Apr 2025", "Post Apr 2025"))) %>%
  mutate(nature = factor(nature, 
                          levels = c("PUBLIC ASSIST", "CHECK WELFARE", "SUICIDAL SUBJECT"),
                          labels = c("PUBLIC ASSIST", "WELFARE", "SUICIDAL")))

# Plot
absorption_plot <- ggplot(absorption_data, aes(x = nature, y = avg_monthly, fill = period)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::comma(round(avg_monthly, 0))),
            position = position_dodge(width = 0.8),
            vjust = -0.5, size = 5.5) +
  scale_y_continuous(limits = c(0, 480),
                     labels = scales::comma) +
  scale_fill_manual(values = c("Pre Apr 2025" = "#032C3C", "Post Apr 2025" = "#BA351A")) +
  labs(
    title = "EPD Monthly Call Volume by Call Type",
    x = NULL,
    y = "Average Monthly Calls",
    fill = NULL
  ) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.position = "bottom",
    legend.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    aspect.ratio = 1
  )

# Save Plot
ggsave("absorption_plot.jpeg", plot = absorption_plot, width = 6, height = 6, dpi = 300)

# Display
absorption_plot
```


### 15.4: DiD for Check Welfare and Suicidal Subject

```{r}
# Per-call-type DiD for Check Welfare and Suicidal Subject
# EPD
epd_calltype_monthly <- eugene %>%
  filter(agency == "EPD") %>%
  filter(nature %in% c("CHECK WELFARE", "SUICIDAL SUBJECT")) %>%
  group_by(year_month, nature, period) %>%
  summarise(call_count = n(), .groups = "drop") %>%
  mutate(city = "Eugene") %>%
  rename(call_type = nature)

# SPD
spd_calltype_monthly <- spd %>%
  filter(agency == "SPD") %>%
  filter(`Final Call Type` %in% c("CHECK WELFARE", "SUICIDAL SUBJECT")) %>%
  group_by(year_month, `Final Call Type`, period) %>%
  summarise(call_count = n(), .groups = "drop") %>%
  mutate(city = "Springfield") %>%
  rename(call_type = `Final Call Type`)

# Combined
calltype_combined <- bind_rows(epd_calltype_monthly, spd_calltype_monthly) %>%
  filter(as.character(year_month) != "2023-01-01") %>%
  filter(as.character(year_month) != "2025-04-01")

calltype_did_key <- calltype_combined %>%
  filter(
    (period == "after") |
    (period == "before" & year_month >= as.Date("2024-04-01") & year_month <= as.Date("2024-12-01"))
  ) %>%
  mutate(
    treat = if_else(city == "Eugene", 1, 0),
    post = if_else(period == "after", 1, 0),
    treat_post = treat * post
  ) %>%
  group_by(call_type) %>%
  do({
    model <- lm(call_count ~ treat + post + treat_post, data = .)
    tibble(
      did_estimate = round(coef(model)["treat_post"], 1),
      p_value = summary(model)$coefficients["treat_post", "Pr(>|t|)"]
    )
  }) %>%
  ungroup()

print(calltype_did_key)
```

### 15.5: Parallel trends for Check Welfare and Suicidal Subject

```{r}
# Parallel trends for Check Welfare and Suicidal Subject
calltype_parallel <- calltype_combined %>%
  filter(call_type %in% c("CHECK WELFARE", "SUICIDAL SUBJECT")) %>%
  filter(period == "before") %>%
  filter(year_month >= as.Date("2024-04-01") & year_month <= as.Date("2024-12-01")) %>%
  mutate(
    treat = if_else(city == "Eugene", 1, 0),
    time_index = as.numeric(year_month)
  ) %>%
  group_by(call_type) %>%
  do({
    model <- lm(call_count ~ treat + time_index + treat:time_index, data = .)
    tibble(
      interaction_estimate = coef(model)["treat:time_index"],
      p_value = summary(model)$coefficients["treat:time_index", "Pr(>|t|)"]
    )
  }) %>%
  ungroup()

print(calltype_parallel)
```

### 15.6: Public Assist t-test

```{r}
# Public Assist t-test because no SPD equivalent
pa_data <- eugene %>%
  filter(agency == "EPD", nature == "PUBLIC ASSIST") %>%
  filter(
    (period == "after") |
    (period == "before" & calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(period, year_month) %>%
  summarise(call_count = n(), .groups = "drop") %>%
  mutate(post = if_else(period == "after", 1, 0))

pa_model <- lm(call_count ~ post, data = pa_data)
summary(pa_model)
```


### 15.7: Eugene Monthly Suicide Calls

```{r}
# Create a table that compares monthly suicide calls before and after by agency

# EPD suicides
epd_suicidal <- eugene %>%
  filter(agency == "EPD", nature == "SUICIDAL SUBJECT") %>%
  filter(
    (period == "after") |
    (period == "before" & calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(period) %>%
  summarise(avg_monthly = round(mean(n), 1), .groups = "drop") %>%
  mutate(agency = "EPD")

# CAHOOTS suicidal (eugene)
cahoots_suicidal <- eugene %>%
  filter(agency == "CAHOOTS", nature == "SUICIDAL SUBJECT") %>%
  filter(
    (period == "after") |
    (period == "before" & calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(period) %>%
  summarise(avg_monthly = round(mean(n), 1), .groups = "drop") %>%
  mutate(agency = "CAHOOTS")

# SPD suicidal
spd_suicidal <- spd %>%
  filter(agency == "SPD", `Final Call Type` == "SUICIDAL SUBJECT") %>%
  filter(
    (period == "after") |
    (period == "before" & `Call Creation Time` >= as.POSIXct("2024-04-01") & 
       `Call Creation Time` <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(period) %>%
  summarise(avg_monthly = round(mean(n), 1), .groups = "drop") %>%
  mutate(agency = "SPD")

# MCSLC suicidal
mcslc_suicidal <- mcslc %>%
  filter(City == "Eugene", `Reason for Dispatch #1` == "Suicidality or suicide attempt") %>%
  filter(
    (period == "after") |
    (period == "before" & `Dispatch Request Date & Time` >= as.POSIXct("2024-04-01") & 
       `Dispatch Request Date & Time` <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(period) %>%
  summarise(avg_monthly = round(mean(n), 1), .groups = "drop") %>%
  mutate(agency = "MCSLC")

# Combine and pivot wide
bind_rows(epd_suicidal, cahoots_suicidal, spd_suicidal, mcslc_suicidal) %>%
  pivot_wider(names_from = period, values_from = avg_monthly, values_fill = 0) %>%
  select(agency, before, after) %>%
  print()
```

### 15.8: EPD Lost Calls Analysis

```{r}
# EPD biggest percentage decreases — matched window, min 40 calls/month
epd_decreases <- eugene %>%
  filter(agency == "EPD") %>%
  filter(
    (period == "after") |
    (period == "before" & calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(nature, period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(nature, period) %>%
  summarise(avg_monthly = round(mean(n), 0), .groups = "drop") %>%
  pivot_wider(names_from = period, values_from = avg_monthly) %>%
  filter(!is.na(before) & !is.na(after)) %>%
  filter(before >= 40) %>%
  mutate(EPD_Pct_Change = round((after - before) / before * 100, 1)) %>%
  filter(EPD_Pct_Change < 0) %>%
  rename(EPD_Before = before, EPD_After = after)

# SPD matched window for comparison
spd_decreases <- spd %>%
  filter(agency == "SPD") %>%
  filter(
    (period == "after") |
    (period == "before" & `Call Creation Time` >= as.POSIXct("2024-04-01") & `Call Creation Time` <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(`Final Call Type`, period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(`Final Call Type`, period) %>%
  summarise(avg_monthly = round(mean(n), 0), .groups = "drop") %>%
  pivot_wider(names_from = period, values_from = avg_monthly) %>%
  mutate(SPD_Pct_Change = round((after - before) / before * 100, 1)) %>%
  rename(nature = `Final Call Type`, SPD_Before = before, SPD_After = after)

# Combine
decreases_table <- epd_decreases %>%
  left_join(spd_decreases, by = "nature") %>%
  select(nature, EPD_Before, EPD_After, EPD_Pct_Change, SPD_Before, SPD_After, SPD_Pct_Change) %>%
  arrange(EPD_Pct_Change)

print(decreases_table, n = 10)
```

### 15.9: EPD Call Reprioritization Chart

```{r}
# Find how certain call types were reprioritized
reprioritization_data <- eugene %>%
  filter(agency == "EPD") %>%
  filter(nature %in% c("CRIMINAL TRESPASS", "BURGLARY", "OVERDOSE", "ILLEGAL BURNING")) %>%
  filter(
    (period == "after") |
    (period == "before" & calltime >= as.POSIXct("2024-04-01") & calltime <= as.POSIXct("2024-12-31"))
  ) %>%
  group_by(nature, period, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(nature, period) %>%
  summarise(avg_monthly = mean(n), .groups = "drop") %>%
  mutate(period = factor(period, levels = c("before", "after"),
                          labels = c("Pre Apr 2025", "Post Apr 2025"))) %>%
  mutate(nature = factor(nature, 
                          levels = c("CRIMINAL TRESPASS", "ILLEGAL BURNING", "BURGLARY", "OVERDOSE"),
                          labels = c("TRESPASS", "ILLEGAL BURN", "BURGLARY", "OVERDOSE")))

reprioritization_plot <- ggplot(reprioritization_data, aes(x = nature, y = avg_monthly, fill = period)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::comma(round(avg_monthly, 0))),
            position = position_dodge(width = 0.8),
            vjust = -0.5, size = 5.5) +
  scale_y_continuous(limits = c(0, 560),
                     labels = scales::comma) +
  scale_fill_manual(values = c("Pre Apr 2025" = "#032C3C", "Post Apr 2025" = "#BA351A")) +
  labs(
    title = "EPD Monthly Call Volume by Call Type",
    x = NULL,
    y = "Average Monthly Calls",
    fill = NULL
  ) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.position = "bottom",
    legend.text = element_text(size = 14),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 14),
    aspect.ratio = 1
  )

reprioritization_plot
ggsave("reprioritization_plot.jpeg", plot = reprioritization_plot, width = 6, height = 6, dpi = 300)
```
