---
title: "CAHOOTS Project"
author: "Alex Gollier"
output: html_document
date: "2026-06-08"
---

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggplot2)
```


## 1. Loading and Merging 

### 1.1 Load and Merge Eugene CAD Data

```{r}
# Set the path to the folder with Eugene CAD csv files
eugene_folder <- "Eugene_CAD_data_noloc"

# Load all of the Eugene CAD data
eugene_files <- list.files(
  # Set path to folder with csv files
  path = eugene_folder,
  # Only grab .csv files
  pattern = "\\.csv$",  
  # Include the folder
  full.names = TRUE
)

# Read every csv file and stack them into one dataframe (also treat NULL values as NA)
eugene_raw <- map_dfr(eugene_files, read_csv, show_col_types = FALSE, na = c("", "NA", "NULL"))

# Look at the data to ensure it was loaded correctly
glimpse(eugene_raw)
```


### 1.2 Load and Merge SPD Data

```{r}
# Set path to SPD file
spd_file <- "2015-2025 SPD Calls for Service.xlsx"

# Get names of all tabs
spd_sheets <- excel_sheets(spd_file)

# Read and stack all tabs, reading all columns as text to avoid type conflicts
spd_raw <- map_dfr(spd_sheets, ~ read_xlsx(spd_file, sheet = .x, col_types = "text"))

cat("SPD rows:", nrow(spd_raw), "\n")  # check row count
glimpse(spd_raw) 
```


### 1.3 Load MCSLC Data

```{r}
# Read the MCSLC file while treating non-city values as NA
mcslc_raw <- read_xlsx("MCSLC.xlsx", na = c("", "Other", "Out of County", "Unknown"))

# Look at the data
glimpse(mcslc_raw)                    
```
## 2. Duplicates

```{r}
# Remove duplicate rows based on unique ID columns
eugene <- eugene_raw %>% distinct(inci_id, .keep_all = TRUE)
spd <- spd_raw %>% distinct(`Incident Number`, .keep_all = TRUE)
mcslc <- mcslc_raw %>% distinct(ID, .keep_all = TRUE)
```

## 3. Missing Values

### 3.1 Eugene Data

```{r}
# Look at total number of rows
cat("Eugene CAD (", nrow(eugene), "rows)\n")

# Check NAs across the important columns in the Eugene dataset
colSums(is.na(eugene[c("inci_id", "calltime", "agency", "nature", "primeunit")]))
```

```{r}
# Drop rows with missing nature, priority, or primeunit
eugene <- eugene %>% filter(!is.na(nature) & !is.na(primeunit))

# Verify rows were dropped
colSums(is.na(eugene[c("inci_id", "calltime", "agency", "nature", "primeunit")]))
```

### 3.2 SPD Data

```{r}
# Look at total number of SPD rows
cat("SPD (", nrow(spd), "rows)\n")

# Count the number of NA values in the relevant variables 
colSums(is.na(spd[c("Incident Number", "Call Creation Time", "Primary Responding Unit", "Final Call Type")]))
```

```{r}
# Drop rows with missing Final Call Type or Priority (only a small percentage of the total dataset)
spd <- spd %>% filter(!is.na(`Final Call Type`) & !is.na(`Priority`) & !is.na(`Primary Responding Unit`))

# Confirm no NAs remain in the important columns
colSums(is.na(spd[c("Incident Number", "Call Creation Time", "Primary Responding Unit", "Final Call Type")]))
```


### 3.3 MCSLC Data

```{r}
# Count total number of MCSLC rows
cat("MCSLC (", nrow(mcslc), "rows)\n")

# Get the number of NA values for the relevant columns
colSums(is.na(mcslc[c("ID", "City", "Dispatch Request Date & Time", "Reason for Dispatch #1")]))
```



```{r}
# Drop rows with missing city, dispatch request date, or reason for dispatch
mcslc <- mcslc %>% filter(!is.na(`Dispatch Request Date & Time`) & !is.na(`Reason for Dispatch #1`) & !is.na(City))

# Confirm no NAs remain
colSums(is.na(mcslc[c("ID", "City", "Dispatch Request Date & Time", "Reason for Dispatch #1")]))
```

## 4. Datetime Variables

```{r}
# Check datetime columns across all three datasets
cat("Eugene calltime:\n")
head(eugene$calltime)

cat("\nSPD Call Creation Time:\n")
head(spd$`Call Creation Time`)

cat("\nMCSLC Dispatch Request Date & Time:\n")
head(mcslc$`Dispatch Request Date & Time`)

# Convert SPD Call Creation Time from Excel serial number to datetime
spd <- spd %>% mutate(`Call Creation Time` = as.POSIXct(
    # Convert days to seconds
    as.numeric(`Call Creation Time`) * 86400,
    # Excel's date system starts from this date
    origin = "1899-12-30",
    # Keep timezone consistent with the other datasets
    tz = "UTC"
  ))

# Confirm conversion worked
cat("\nSPD Call Creation Time Fixed:\n")
head(spd$`Call Creation Time`)
```

## 5. Agency Variables


### 5.1 Eugene Data

```{r}
# Label CAHOOTS calls using the given format patterns, default everything else to EPD
eugene <- eugene %>%
  mutate(agency = case_when(
    agency == "CAHE" ~ "CAHOOTS",
    !is.na(primeunit) & primeunit == "_CAHOT" ~ "CAHOOTS",
    !is.na(primeunit) & str_detect(primeunit, "^_\\dJ\\d{2}$") ~ "CAHOOTS",
    !is.na(nature) & str_detect(nature, "CAHOOTS") ~ "CAHOOTS",
    TRUE ~ "EPD"
  ))

# Clean nature column by stripping the CAHOOTS suffix     
eugene <- eugene %>%
  mutate(nature = str_remove(nature, ",\\s*CAHOOTS.*$")) %>%
  mutate(nature = if_else(str_detect(nature, "PUBLIC ASSIST|ASSIST PUBLIC"), "PUBLIC ASSIST", nature))

# Confirm labeling worked
eugene %>% count(agency)
```

### 5.2 SPD Data

```{r}
# Create an agency column that labels CAHOOTS calls using the #J## pattern 
spd <- spd %>%
  mutate(agency = if_else(
    # Use a regex pattern
    str_detect(`Primary Responding Unit`, "^\\dJ\\d{2}$"),  
    "CAHOOTS",
    "SPD"
  ))

# Confirm labeling worked
spd %>% count(agency)
```   

### 5.3 MCSLC Data

```{r}
# All rows are MCSLC so just add a label row
mcslc <- mcslc %>% mutate(agency = "MCSLC")

# Confirm it worked
mcslc %>% count(agency)
```

## 6. Period Variable

```{r}
# Add period variable to Eugene CAD 
eugene <- eugene %>% mutate(period = if_else(calltime < as.POSIXct("2025-04-08"), "before", "after"))

# Add period variable to SPD
spd <- spd %>% mutate(period = if_else(`Call Creation Time` < as.POSIXct("2025-04-08"), "before", "after"))

# Add period variable to MCSLC
mcslc <- mcslc %>% mutate(period = if_else(`Dispatch Request Date & Time` < as.POSIXct("2025-04-08"), "before", "after"))

# Confirm period variable in each dataset
eugene %>% count(period)
spd %>% count(period)
mcslc %>% count(period)
```
## 7. Month Variable

```{r}
# Add year-month column to each dataset

# EPD
eugene <- eugene %>% mutate(year_month = floor_date(calltime, "month"))

# SPD
spd <- spd %>% mutate(year_month = floor_date(`Call Creation Time`, "month"))

# MCSLC
mcslc <- mcslc %>% mutate(year_month = floor_date(`Dispatch Request Date & Time`, "month"))
```


## 8. Select Columns

```{r}
# Keep only relevant columns in Eugene CAD
eugene <- eugene %>% select(inci_id, calltime, agency, nature, period, year_month, primeunit)

# Keep only relevant columns in SPD
spd <- spd %>% select(`Incident Number`, `Call Creation Time`, `Primary Responding Unit`, `Final Call Type`, agency, period, year_month)

# Keep only relevant columns in MCSLC
mcslc <- mcslc %>% select(ID, City, `Dispatch Request Date & Time`, `Reason for Dispatch #1`, agency, period, year_month)

# Confirm columns in each dataset
head(eugene)
head(spd)
head(mcslc)
```

## 9. Monthly Aggregation

```{r}
# Find monthly call counts by agency and period

# EPD
eugene_monthly <- eugene %>%
  group_by(year_month, agency, period) %>%
  summarise(call_count = n(), .groups = "drop") %>%
  mutate(city = "Eugene")

# SPD
spd_monthly <- spd %>%
  group_by(year_month, agency, period) %>%
  summarise(call_count = n(), .groups = "drop") %>%
  mutate(city = "Springfield")

# MCSLC
mcslc_monthly <- mcslc %>%
  group_by(year_month, City, agency, period) %>%
  summarise(call_count = n(), .groups = "drop") %>%
  rename(city = City)

# Combine into one dataframe
monthly_combined <- bind_rows(eugene_monthly, spd_monthly, mcslc_monthly) %>%
  # Drop January 2023 due to weird missing values 
  filter(as.character(year_month) != "2023-01-01") %>%    
  # Drop April 2025 due to split
  filter(as.character(year_month) != "2025-04-01")

# Check structure of results
glimpse(eugene_monthly)
glimpse(spd_monthly)
glimpse(mcslc_monthly)
glimpse(monthly_combined)
```

## 10. Proportion Calculation

```{r}
# Calculate total calls per city-month and each agency's proportion
monthly_combined <- monthly_combined %>%
  # Group by city and month
  group_by(year_month, city) %>%
  # Find proportions and get rid of extra grouping
  mutate(total_calls = sum(call_count), proportion = call_count / total_calls) %>% ungroup()
```

# Save cleaned objects for analysis
saveRDS(eugene, "eugene_cleaned.rds")
saveRDS(spd, "spd_cleaned.rds")
saveRDS(mcslc, "mcslc_cleaned.rds")
saveRDS(monthly_combined, "monthly_combined.rds")
